---
title: "mis_indep_of_Y"
output: html_document
date: '2022-07-26'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(mice))
suppressPackageStartupMessages(library(tidyverse))
```

## 2) Probability of Missing data doesn't depend on Response Y 
  * Similarly to above, suppose scientifc interest focuses on determining $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$ in the linear model $y_i = \alpha + x_1\beta_1 + x_2\beta_2 + x_3\beta_3 + x_4\beta_4 + \epsilon_i$. 
  * Here, $\epsilon_i ~ N(0, \sigma^2)$.
  * The missingness model gives observations with X_2 > median(X_2) a different probability of missingness in X_1, X_3, X_4 than for observations with X_2 <= median(X_2) 

```{r}
# Draw data from artificial model specified above
create.data <- function(alpha = 20, beta_1 = 1, beta_2 = 2, beta_3 = 3, beta_4 = 4, 
                        sigma2 = 1, n = 200, run = 1) {
  set.seed(seed = run)
  x_1 <- rnorm(n)
  x_2 <- rnorm(n)
  x_3 <- rnorm(n)
  x_4 <- rnorm(n)
  y <- beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + alpha + rnorm(n, sd = sqrt(sigma2))
  as.data.frame(cbind("Y" = y, "X_1"= x_1, "X_2" = x_2, "X_3" = x_3, "X_4" = x_4))
  
}
```

```{r}
data <- create.data()

head(data)
```

```{r}

MNAR.make.missing <- function(data, prob_missing_larger = 0.2, 
                              prob_missing_smaller = 0.8){
  # Setting up the randomness categories for missingness in X_1, X_3, X_4 
  higher <- data %>% filter(X_2 > median(X_2)) %>% select(X_1) 
  rx1_larger <- rbinom(nrow(higher), 1, prob_missing_larger)
  rx1_smaller <- rbinom(nrow(data) - nrow(higher), 1, prob_missing_smaller)
  rx3_larger <- rbinom(nrow(higher), 1, prob_missing_larger)
  rx3_smaller <- rbinom(nrow(data) - nrow(higher), 1, prob_missing_smaller)
  rx4_larger <- rbinom(nrow(higher), 1, prob_missing_larger)
  rx4_smaller <- rbinom(nrow(data) - nrow(higher), 1, prob_missing_smaller)
  rx1 <- c(rx1_larger, rx1_smaller)
  rx3 <- c(rx3_larger, rx3_smaller)
  rx4 <- c(rx4_larger, rx4_smaller)
  data <- data %>% 
    arrange(desc(X_2)) %>%
    cbind(rx1, rx3, rx4)
  # Implementing the missingness in X_1, X_3, X_4
  data <- data %>% mutate(X_1 = case_when(rx1 == 1 ~ as.numeric(NA), 
                    rx1 == 0 ~ as.numeric(data$X_1))) %>%
    mutate(X_3 = case_when(rx3 == 1 ~ as.numeric(NA), 
                    rx3 == 0 ~ as.numeric(data$X_3))) %>% 
    mutate(X_4 = case_when(rx4 == 1 ~ as.numeric(NA), 
                    rx4 == 0 ~ as.numeric(data$X_4)))
  # Remove setup variables
  data <- select(data, -c(rx1, rx3, rx4)) 
  data
}


```

```{r}

data_mis <- MNAR.make.missing(data, 0.2, 0.8)

data_mis
```
### Multiple Imputation

```{r}
# Simulate multiple imputation, obtaining estimates and 95% confidence interval.
simulate_MI2 <- function(runs = 100) {
  res <- array(NA, dim = c(5, runs, 3))
  times <- array(NA, dim = c(100, 1, 1))
  dimnames(res) <- list(c("Intercept", "X_1", "X_2", "X_3", "X_4"),
                        as.character(1:runs), c("estimate", "2.5%", "97.5%"))
  sim_dataset <- as.data.frame(create.data(n = 200))
  for (run in 1:runs){
      # Note that time is only measured for the MI/imp steps 
      # (i.e. filtering, predicting)
    missingness_sim_dataset <- MNAR.make.missing(sim_dataset, 0.2, 0.8)
    start_time <- Sys.time()
    imp_MI <- mice(missingness_sim_dataset, print = FALSE)
    fit <- with(imp_MI, lm(Y ~ X_1 + X_2 + X_3 + X_4))
    end_time <- Sys.time()
    tab <- summary(pool(fit), "all", conf.int = TRUE)  
    res[1, run, ] <- as.numeric(tab[1, c("estimate", "2.5 %", "97.5 %")])  
    res[2, run, ] <- as.numeric(tab[2, c("estimate", "2.5 %", "97.5 %")])
    res[3, run, ] <- as.numeric(tab[3, c("estimate", "2.5 %", "97.5 %")])
    res[4, run, ] <- as.numeric(tab[4, c("estimate", "2.5 %", "97.5 %")])
    res[5, run, ] <- as.numeric(tab[5, c("estimate", "2.5 %", "97.5 %")])
    
    times[run, 1, 1] <- as.numeric(end_time - start_time)
  }
  list(res, times)
}

```

```{r}
# Run 100 iterations 
res_MI2 <- simulate_MI2(100)

```

```{r}
# Obtain confidence intervals & estimates for all coefficients, intercept.
apply(res_MI2[[1]], c(1, 3), mean, na.rm = TRUE)

```

```{r}
# Mean time for the multiple imputation instances
times <- res_MI2[[2]]
mean(times)
```



```{r}
# Evaluating imputation method performance for estimating
# all parameters of interest.
res <- res_MI2[[1]]
true <- c(20, 1, 2, 3, 4)
RB <- rowMeans(res[,, "estimate"]) - true
PB <- 100 * abs((rowMeans(res[,, "estimate"]) - true)/ true)
CR <- rowMeans(res[,, "2.5%"] < true & true < res[,, "97.5%"])
AW <- rowMeans(res[,, "97.5%"] - res[,, "2.5%"])
RMSE <- sqrt(rowMeans((res[,, "estimate"] - true)^2))
data.frame(RB, PB, CR, AW, RMSE)
```

###  Listwise Deletion

```{r}
# Simulate listwise deletion, obtaining estimates and 95% confidence interval.

simulate_LD <- function(runs = 100){
  res <- array(NA, dim = c(5, runs, 3))
  dimnames(res) <- list(c("Intercept", "X_1", "X_2", "X_3", "X_4"),
                        as.character(1:runs), c("estimate", "2.5%", "97.5%"))
  times <- array(NA, dim = c(runs, 1, 1))
  sim_dataset <- as.data.frame(create.data(n = 200))
  # Note that time is only measured for the LD/imp steps (i.e. filtering, predicting)
  for (run in 1:runs){
    missingness_sim_dataset <- MNAR.make.missing(sim_dataset, 0.2, 0.8)
    start_time <- Sys.time()
    filtered_sim_dataset <- missingness_sim_dataset %>% 
      select(Y, X_1, X_2, X_3, X_4) %>%
      filter(!is.na(X_1), !is.na(X_3), !is.na(X_4))
    fit <- with(filtered_sim_dataset, lm(Y ~ X_1 + X_2 + X_3 + X_4))
    end_time <- Sys.time()
    times[run, 1, 1] <- as.numeric(end_time - start_time)
    # loop over each variable. Note we do the imputation just ONCE b/c LD is 
    # deterministic.
    for (var in 1:5){
      edges <- as.numeric((confint(fit)[var,]))
      estimate <- as.numeric(fit$coefficients)[var]
      interval <- c(estimate, edges)
      res[var, run, ] <- interval
      }
  
  }
  list(res, times)
}

```

```{r}
result_LD <- simulate_LD()

```

```{r}
# Obtain confidence intervals & estimates for all coefficients, intercept.
apply(result_LD[[1]], c(1, 3), mean, na.rm = TRUE)

```

```{r}
# Evaluating imputation method performance for estimating
# all parameters of interest.
res <- result_LD[[1]]
true <- c(20, 1, 2, 3, 4)
RB <- rowMeans(res[,, "estimate"]) - true
PB <- 100 * abs((rowMeans(res[,, "estimate"]) - true)/ true)
CR <- rowMeans(res[,, "2.5%"] < true & true < res[,, "97.5%"])
AW <- rowMeans(res[,, "97.5%"] - res[,, "2.5%"])
RMSE <- sqrt(rowMeans((res[,, "estimate"] - true)^2))
data.frame(RB, PB, CR, AW, RMSE)
```

```{r}
# Mean time for 100 instances of LD
times_LD <- result_LD[[2]]
mean(times_LD)
```

