---
title: "Multiple Imputation Edge Cases"
output:
  pdf_document: default
  html_document: default

date: "2022-07-22, Leo Murao Watson"
---
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(mice))
suppressPackageStartupMessages(library(tidyverse))
suppressPackageStartupMessages(library(kableExtra))
suppressPackageStartupMessages(library(dplyr))
```

## Special Cases where Listwise Deletion is Preferred over Multiple Imputation
  * Let $Y$ = Ozone, $X_1$ = Wind, $X_2$ = Temp, $X_3$ = Month, $X_4$ = Day
  * Will compare Missing Imputation and Listwise Deletion as missing data methods.
  * Suppose scientifc interest focuses on determining $\beta_1$, $\beta_2$, $\beta_3$, $\beta_4$ in the linear model $y_i = \alpha + x_1\beta_1 + x_2\beta_2 + x_3\beta_3 + x_4\beta_4 + \epsilon_i$. 
  *Here, $\epsilon_i ~ N(0, \sigma^2)$.
  
## 1) Exclusively Missing data in Response Y 
  
  
```{r}
# Draw data from artificial model specified above
create.data <- function(alpha = 20, beta_1 = 1, beta_2 = 2, beta_3 = 3, beta_4 = 4, 
                        sigma2 = 1, n = 50, run = 1) {
  set.seed(seed = run)
  x_1 <- rnorm(n)
  x_2 <- rnorm(n)
  x_3 <- rnorm(n)
  x_4 <- rnorm(n)
  y <- beta_1 * x_1 + beta_2 * x_2 + beta_3 * x_3 + beta_4 * x_4 + alpha + 
    rnorm(n, sd = sqrt(sigma2))
  cbind("Y" = y, "X_1"= x_1, "X_2" = x_2, "X_3" = x_3, "X_4" = x_4)
}

# Remove some data to simulate real world missingness
MCAR.make.missing <- function(data, p = 0.5){
  rx <- rbinom(nrow(data), 1, p)
  data[rx == 0, "Y"] <- NA
  data
}
```



### Missing Imputation


```{r}
# Simulate multiple imputation, obtaining estimates and 95% confidence interval.
simulate_MI2 <- function(runs = 100) {
  res <- array(NA, dim = c(5, runs, 3))
  times <- array(NA, dim = c(100, 1, 1))
  dimnames(res) <- list(c("Intercept", "X_1", "X_2", "X_3", "X_4"),
                        as.character(1:runs), c("estimate", "2.5%", "97.5%"))
  sim_dataset <- as.data.frame(create.data(n = 200))
  for (run in 1:runs){
      # Note that time is only measured for the MI/imp steps 
      # (i.e. filtering, predicting)
    missingness_sim_dataset <- MCAR.make.missing(sim_dataset, p = 0.5)
    start_time <- Sys.time()
    imp_MI <- mice(missingness_sim_dataset, print = FALSE)
    fit <- with(imp_MI, lm(Y ~ X_1 + X_2 + X_3 + X_4))
    end_time <- Sys.time()
    tab <- summary(pool(fit), "all", conf.int = TRUE)  
    res[1, run, ] <- as.numeric(tab[1, c("estimate", "2.5 %", "97.5 %")])  
    res[2, run, ] <- as.numeric(tab[2, c("estimate", "2.5 %", "97.5 %")])
    res[3, run, ] <- as.numeric(tab[3, c("estimate", "2.5 %", "97.5 %")])
    res[4, run, ] <- as.numeric(tab[4, c("estimate", "2.5 %", "97.5 %")])
    res[5, run, ] <- as.numeric(tab[5, c("estimate", "2.5 %", "97.5 %")])
    
    times[run, 1, 1] <- as.numeric(end_time - start_time)
  }
  list(res, times)
}

```

```{r}
# Run 100 iterations 
res_MI2 <- simulate_MI2(100)
```


```{r}
# Obtain confidence intervals & estimates for all coefficients, intercept.
apply(res_MI2[[1]], c(1, 3), mean, na.rm = TRUE)

```

```{r}
# Mean time for the multiple imputation instances
times <- res_MI2[[2]]
mean(times)
Multiple_Imputation <- mean(times)
```

```{r}
# Evaluating imputation method performance for estimating
# all parameters of interest.
res <- res_MI2[[1]]
true <- c(20, 1, 2, 3, 4)
Raw_Bias <- rowMeans(res[,, "estimate"]) - true
Percent_Bias <- 100 * abs((rowMeans(res[,, "estimate"]) - true)/ true)
Coverage_Rate <- rowMeans(res[,, "2.5%"] < true & true < res[,, "97.5%"])
Avg_Width <- rowMeans(res[,, "97.5%"] - res[,, "2.5%"])
RMSE <- sqrt(rowMeans((res[,, "estimate"] - true)^2))
MI_measures <- data.frame(Percent_Bias, Coverage_Rate, Avg_Width, RMSE)
knitr::kable(round(MI_measures, 3), align = "ccccc") %>% kable_styling(full_width = F)

```


###  Listwise Deletion


```{r}
# Simulate listwise deletion, obtaining estimates and 95% confidence interval.

simulate_LD <- function(runs = 100){
  res <- array(NA, dim = c(5, runs, 3))
  dimnames(res) <- list(c("Intercept", "X_1", "X_2", "X_3", "X_4"),
                        as.character(1:runs), c("estimate", "2.5%", "97.5%"))
  times <- array(NA, dim = c(runs, 1, 1))
  sim_dataset <- as.data.frame(create.data(n = 200))
  # Note that time is only measured for the LD/imp steps (i.e. filtering, predicting)
  for (run in 1:runs){
    missingness_sim_dataset <- MCAR.make.missing(sim_dataset, p = 0.7)
    start_time <- Sys.time()
    filtered_sim_dataset <- missingness_sim_dataset %>% 
      select(Y, X_1, X_2, X_3, X_4) %>%
      filter(!is.na(Y))
    fit <- with(filtered_sim_dataset, lm(Y ~ X_1 + X_2 + X_3 + X_4))
    end_time <- Sys.time()
    times[run, 1, 1] <- as.numeric(end_time - start_time)
    # loop over each variable. Note we do the imputation just ONCE b/c LD is 
    # deterministic.
    for (var in 1:5){
      edges <- as.numeric((confint(fit)[var,]))
      estimate <- as.numeric(fit$coefficients)[var]
      interval <- c(estimate, edges)
      res[var, run, ] <- interval
      }
  
  }
  list(res, times)
}

```

```{r}
result_LD <- simulate_LD()
```




```{r}
# Obtain confidence intervals & estimates for all coefficients, intercept.
apply(result_LD[[1]], c(1, 3), mean, na.rm = TRUE)

```

```{r}
# Evaluating imputation method performance for estimating
# all parameters of interest.
res <- result_LD[[1]]
true <- c(20, 1, 2, 3, 4)
Raw_Bias <- rowMeans(res[,, "estimate"]) - true
Percent_Bias <- 100 * abs((rowMeans(res[,, "estimate"]) - true)/ true)
Coverage_Rate <- rowMeans(res[,, "2.5%"] < true & true < res[,, "97.5%"])
Avg_Width <- rowMeans(res[,, "97.5%"] - res[,, "2.5%"])
RMSE <- sqrt(rowMeans((res[,, "estimate"] - true)^2))
LD_measures <- data.frame(Percent_Bias, Coverage_Rate, Avg_Width, RMSE)
```
```{r}
knitr::kable(round(LD_measures, 3), align = "ccccc") %>% kable_styling(full_width = F)

```

```{r}
round(LD_measures, 3)
```

```{r}
# Mean time for 100 instances of LD
times_LD <- result_LD[[2]]
ListwiseDeletion <- mean(times_LD)
```
```{r}
mean_times <- as.data.frame(rbind(Multiple_Imputation, ListwiseDeletion), col.names = "average_runtime")
colnames(mean_times) <- "average-runtime"
mean_times
knitr::kable(mean_times) %>% kable_styling(full_width = F)
```
```{r}
```


<!-- ##### Highlight and command shift C to uncomment below -->
<!-- ##### Testing -->

<!-- ```{r} -->

<!-- airquality_processed <- airquality %>% select(Ozone, Wind, Temp, Month, Day) -->

<!-- imp_MI <- mice(airquality_processed, print = FALSE) -->
<!-- fit <- with(imp_MI, lm(Ozone ~ Wind + Temp + Month + Day)) -->
<!-- tab <- summary(pool(fit), "all", conf.int = TRUE) -->
<!-- tab -->


<!-- # intercept_interval <- as.numeric(tab[1, c("estimate", "2.5 %", "97.5 %")]) -->
<!-- # wind_coeff_interval <- as.numeric(tab[2, c("estimate", "2.5 %", "97.5 %")]) -->
<!-- # temp_coeff_interval <- as.numeric(tab[3, c("estimate", "2.5 %", "97.5 %")]) -->
<!-- # month_coeff_interval <- as.numeric(tab[4, c("estimate", "2.5 %", "97.5 %")]) -->
<!-- # day_coeff_interval <- as.numeric(tab[5, c("estimate", "2.5 %", "97.5 %")]) -->
<!-- # typeof(day_coeff_interval) -->
<!-- ``` -->

<!-- ```{r} -->

<!-- ```{r} -->
<!-- fit$coefficients -->
<!-- typeof(fit$coefficients) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- confint(fit)  -->
<!-- typeof(confint(fit)) -->
<!-- ``` -->
<!-- ```{r} -->
<!-- # Intercept case -->
<!-- edges <- as.numeric((confint(fit)[1,])) -->
<!-- mid <- as.numeric(fit$coefficients)[1] -->
<!-- interval <- c(edges[1], mid, edges[2]) -->
<!-- ``` -->


<!-- ```{r} -->
<!-- res <- array(NA, dim = c(5, 1, 3)) -->

<!-- ``` -->


