---
title: "BOCF"
output: html_document
date: '2022-06-21'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(mice))
suppressPackageStartupMessages(library(tidyverse))

airquality_indicator
```

# BOCF (longitudinal data imputation)

### Description
* BOCF (Base Observation Carried Forward) is a way of imputing longitudinal data.
* As per its name suggests, BOCF imputes missing values by replacing them with some predetermined baseline value. Typically used in clinical trials where a "baseline value" can easily be established. 
  
#### Example
```{r}
# Store baseline (first observation) of relevant variable [Ozone in this case] 
baseline_Ozone = (airquality$Ozone)[1]

# Impute missing values with baseline observation as well as adding ozone missingness indicator variable.
airquality_LOCF <- airquality %>% 
  mutate(imputation_indicator = if_else(is.na(Ozone), "imputed", "observed")) %>% 
  replace_na(list(Ozone = baseline_Ozone, Solar.R = baseline_Solar.R))


```
  
* Notice in BOCF imputed values are equal to the baseline value observed on Day 1
```{r}
# Create line graph of ozone vs day for the first 30 observations, where imputed values are highlighted using red.
head(airquality_LOCF, 30) %>% ggplot(aes(x = Day,y= Ozone)) + 
  geom_line() + 
  geom_point( aes(color = imputation_indicator)) + 
  labs(y = "ozone  (ppb)",
       x = "Day Number")
```

### Pros

* Conceptually very simple in generating a complete dataset.
* Computationally cheap relative to other techniques that may be more statistically sound (e.g. multiple imputation).


### Cons
* Should only be applied when we are confident in cases where we are certain what the missing values should be.
* Can yield biased estimators even under MCAR 
* Panel on Handling Missing Data in Clinical Trials recommends BOCF shouldn't be used unless under very specific circumstances where BOCF assumptions are justified.