---
title: "missing_indicator_method"
output: html_document
date: '2022-06-21'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Missing Indicator Method


### Description

* Imputation technique applicable when explanatory variable(s) have missingness but response does not.
* Missing values are replaced by a fixed value (usually 0 or observed values’ mean), and a missingness indicator variable is added to each observation in the dataset for each incomplete variable.

#### Example
```{r}
airquality
```

```{r}
# Impute the airquality dataset, using observed mean as the fixed imputation value.
imp <- mice(airquality, method = "mean", m = 1,
            maxit = 1, print = FALSE)
# add column to airquality dataset indicating whether Ozone value is missing or not.
airquality2 <- cbind(complete(imp),
                     r.Ozone = is.na(airquality[, "Ozone"]))
# apply multiple linear regression
fit <- lm(Wind ~ Ozone + r.Ozone, data = airquality2)
```
### Pros
* In randomized trials, can be used to generate unbiased estimators in MCAR and MAR covariate situations.
* Retains the whole observed dataset.


### Cons
* In nonrandomized studies (e.g. observational studies), this method can result in biased estimators and associations between variables even in MCAR situations. 
* Does not allow for missingness in response variable.


