---
title: "LOCF"
output: html_document
date: '2022-06-21'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(mice))
suppressPackageStartupMessages(library(tidyverse))
```

# LOCF (longitudinal data imputation)

### Description
* LOCF (Last Observation Carried Forward) is a way of imputing longitudinal data.
* It imputes missing values by replacing them with the latest observed value. Typically seen in clinical trials
  
#### Example
```{r}
airquality_filled <- tidyr::fill(airquality, Ozone)
airquality_filled %>% ggplot(aes(x = Day,y= Ozone)) + 
  geom_point() + 
  labs(y = "ozone  (ppb)",
       x = "Day Number")
```

```{r}
airquality_indicator <- airquality %>% 
  mutate(imputation_indicator = if_else(is.na(Ozone), "imputed", "observed"))

airquality2 <- tidyr::fill(airquality_indicator, Ozone)
head(airquality2, 30) %>% ggplot(aes(x = Day,y= Ozone, color = imputation_indicator)) + 
  geom_line() + 
  geom_point() + 
  labs(y = "ozone  (ppb)",
       x = "Day Number")
```

### Pros

* Conceptually very simple to generate a complete dataset.
* Computationally cheap relative to other techniques that may be more statistically sound (e.g. multiple imputation).


### Cons
* Should only be applied when we are confident in cases where we are certain what the missing values should be.
* Can yield biased estimators even under MCAR 


