---
title: "pairwise_deletion"
output: html_document
date: '2022-06-19'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pairwise Deletion


### Description
* Pairwise Deletion seeks to utilize ALL available data 
* Calculates means of variables using observed data for each variable.
* Combines this with correlations/correlations between the different variables in the dataset to impute values.

#### Example  

* Inspect Dataset
```{r}
suppressPackageStartupMessages(library(lavaan))
library(lavaan)
data <- airquality[, c("Ozone", "Solar.R", "Wind")]
head(data)
```
* Compute means for each column using observed values
```{r}
mu <- colMeans(data, na.rm = TRUE)
mu
```
* Compute covariances for each variable
```{r}
cv <- cov(data, use = "pairwise")
cv
```

Apply pairwise deletion using the Lavaan package to obtain coefficients for Ozone w.r.t Wind, Solar.R in regression model.
```{r cars}
fit <- lavaan("Ozone ~ 1 + Wind + Solar.R
              Ozone ~~ Ozone",
             sample.mean = mu, sample.cov = cv,
             sample.nobs = sum(complete.cases(data)))
coef(fit)[c(2:3)]
```

### Pros
* Pairwise deletion works well if data $(Y, X_1,...,X_n)$ approximately follows a multivariate normal distribution and the correlations between variables are low.
* If MCAR is plausible, provides unbiased estimates effectively and efficiently.


### Cons  
* Requires MCAR assumption for unbiased estimates.
* Correlation matrix obtained via pairwise deletion may give mathematically impossible results, and hence unable to perform regression analysis [see code below]
  + Consider 4x3 data frame with some missingness
```{r}
df <- data.frame(X1 = c(3,NA,2,4),
                 X2 = c(4,1,2,NA),
                 X3 = c(4,1,NA,3))
df
```

  + Create correlation matrix. Note that $X_1, X_2$ are perfectly correlated, $X_2, X_3$ are perfectly correlated and YET $X_1, X_3$ are perfectly negatively correlated. This is a physically impossible scenario.
```{r}
cor(df, use=  "pairwise.complete.obs")
```

* Covariance/Correlation matrix may not be positive definite (requirement for multivariable procedures and further analysis, e.g. MANOVA [multivariate analysis of variance]) 
* Sample size will almost certainly vary between variables due to missingness. This is problematic for calculating standard errors as the $n$ value is ambiguous and hence either underestimated or overestimated.

