---
title: "pairwise_deletion"
output: html_document
date: '2022-06-19'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Pairwise Deletion


### Description
* Pairwise Deletion seeks to utilize ALL available data 
* Calculates means of variables using observed data for each variable.
* Combines this with correlations/correlations between the different variables in the dataset to impute values.

### Example  

Inspect Dataset
```{r}
suppressPackageStartupMessages(library(lavaan))
library(lavaan)
data <- airquality[, c("Ozone", "Solar.R", "Wind")]
head(data)
```
Compute means for each column using observed values
```{r}
mu <- colMeans(data, na.rm = TRUE)
mu
```
Compute covariances for each variable
```{r}
cv <- cov(data, use = "pairwise")
cv
```

Apply pairwise deletion using the Lavaan package
```{r cars}
fit <- lavaan("Ozone ~ 1 + Wind + Solar.R
              Ozone ~~ Ozone",
             sample.mean = mu, sample.cov = cv,
             sample.nobs = sum(complete.cases(data)))
#summary(fit)
# Gives coefficients for regression model.
coef(fit)[c(2:3)]
```

### Pros
* If data $(Y, X_1,...,X_n)$ follows multivariate normal distribution, correlations between variables are low, and if MCAR is plausible, provided unbiased estimates effectively and efficiently.


### Cons
