---
title: "regression_imputation"
output: html_document
date: '2022-06-20'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
suppressPackageStartupMessages(library(mice))
suppressPackageStartupMessages(library(tidyverse))
library(mice)
library(tidyverse)
```

# (Deterministic) Regression Imputation


### Description
* Fits a regression model between observed values of predictor variables $(X_1, ..., X_n)$ and response variable $Y$.
* Imputes NA values with the exact predicted values from regression model.

#### Example
```{r}
# Fit linear regression model to Ozone w.r.t Solar.R from airquality built-in dataset.
fit <- lm(Ozone ~ Solar.R, data = airquality)
# Applies fit to dataset.
pred <- predict(fit, airquality %>% select(Ozone, Solar.R))
pred_df <- as.data.frame(pred)

```
 
```{r}
# Create new variable "model_ozone" that takes exact value given the fitted model, and new variable "combined_ozone" that takes observed "Ozone" value if it's not missing, "model_ozone" value if missing.
airquality_processed <- airquality %>% 
  select(Ozone, Solar.R) %>%
  mutate(model_ozone = pred_df$pred) %>%
  mutate(combined_ozone = if_else(is.na(Ozone), model_ozone, as.numeric(Ozone), missing = NULL)) %>% mutate(combined_ozone_type = if_else(is.na(Ozone), "imputed", "observed", missing = NULL)) %>% relocate(Solar.R, .after = combined_ozone) %>% relocate(combined_ozone)
head(airquality_processed, n = 25)
```
* Note the imputed values in the scatter plot below are exactly the predicted values from the regression fit with no added randomness. This increases the correlation and decreases variance, as shown in the next chunks. 
```{r warning = FALSE}
airquality_processed %>% 
  ggplot(aes(x = Solar.R,y= combined_ozone, color = combined_ozone_type)) + 
  geom_point() + 
  labs(y = "ozone  (ppb)",
       x = "Solar Radiation")
```
```{r}
# Correlation for observed values
cor.test(airquality_processed$Ozone, airquality_processed$Solar.R)$estimate
# Correlation for observed/imputed values 
cor.test(airquality_processed$combined_ozone, airquality_processed$Solar.R)$estimate
```
```{r}
# Variance for observed values
var(airquality_processed$Ozone, airquality_processed$Solar.R, na.rm = TRUE)
# Variance for observed/imputed values
var(airquality_processed$combined_ozone, airquality_processed$Solar.R, na.rm = TRUE)

```




### Pros
* Yields unbiased mean estimates under MCAR, and regression weights of imputation model if the explanatory variables are complete.
* Yields unbiased estimators for regression weights under MAR, if the variable causing the MAR is part of the regression model. (e.g. if Solar.R was responsible for missingness in Ozone)


### Cons
* Since imputed values are the exact predicted values from the regression model, correlation is overestimated and variability is underestimaed
* Imputations are realistic if predicted linear model is almost perfect, but deterministic regression often too precise and too good to be true; spurious relations and false positives abound. 